Matroids \& testing in families of dependent hypotheses

### 2013-08-19 presentation/presentation.tex
cleaned up the first couple of slides

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{ulem}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\em i.e.},~}
\newcommand{\eg}{{\em e.g.},~}
\newcommand{\parentd}{../}
\newcommand{\R}{\mathbb{R}}
\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage

\title{Matroids from families of null hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

@@

\section{Logically dependent hypotheses}

\begin{frame}
\frametitle{Pairwise comparisons, some history }

Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):
\begin{itemize}
  \item Parameters: $\mu_1,\mu_2,\mu_3$ ; \\ Hypotheses: $H_1:\mu_1=\mu_2,H_2:\mu_2=\mu_3;H_3:\mu_1=\mu_3$
  \item A level-$\alpha$ test of $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 : \mu_1=\mu_2=\mu_3$. followed by level-$\alpha$ tests of $H_1,H_2,H_3$ protects the FWER at level $\alpha$.
  \begin{itemize}
    \item  essentially closed test; no need to test pairwise intersection hypotheses since $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j$,
    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Pairwise comparisons, some history}

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
       \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{itemize}

\item Consider $\mu_i, i \in \{1,...,m\}$

\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
\begin{itemize}
\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
\item edges correspond to hypotheses, \emph{i.e.}, $E =
      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
      j$, and $e_l \neq e_k$ for $l \neq k$
\end{itemize}
\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
  \in I\}$
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 1: 4 Groups}
\input{full4ext}
\end{frame}

@@

\begin{tabular}{ll}
  \begin{minipage}{.45\textwidth}
    All pairwise comparisons of 4 means 
    \begin{displaymath}
    \bs{C}^t = \left(
      \begin{array}{llll}
      -1 & 1 & 0 & 0 \\
      -1 & 0 & 1 & 0 \\
      -1 & 0 & 0 & 1 \\
      0 & -1 & 1 & 0 \\
      0 & -1 & 0 & 1 \\
      0 & 0 & -1 & 1 
    \end{array}
    \right)
    \end{displaymath}
  \end{minipage} &
  \begin{minipage}{.55\textwidth}
    Modified Williams Contrasts for Dose Response
    \begin{displaymath}
      \bs{C}^t=\left(
      \begin{array}{llll}
        1 & 0 & 0 & -1\\
        1 & 0 &-.51 & -.49 \\
        1 & -.35 & -.33 & -.32\\
        .51 & .49 & -.51 & -.48\\
        .51 & .49 & 0 &  -1\\
        .35 & .33 & .32& -1
      \end{array}\right)
    \end{displaymath}
  \end{minipage}
\end{tabular}

@@

\begin{itemize}
\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
  dependence relation $\sim_H$ on $\mathcal{P}(S)$
\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq
  H_{I\setminus i}$ for all $i \in I$
\pause
\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
  column rank
\end{itemize} % ends low level
\end{frame}

@@

\item Matroids generalize the structures of graphs and vector spaces,
  \eg cycles, bases, independent sets
\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
  structure 
\item Cryptomorphism: many ways to define a matroid: 
  \begin{itemize}
  \item From a graph perspective: Circuit axioms.
  \item From a vector space perspective: Independence axioms.
  \end{itemize}
\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
\end{itemize} % ends low level
\end{frame}

@@

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \only<2->{\input{cycle}}
.
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
    \begin{enumerate}
    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
      $\exists C_3 \subseteq (C_1 \cup C_2)$,
    \end{enumerate}
    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
  \end{minipage}
\end{tabular}

@@

% \item \textbf{/Some illustration hereabouts/}

% \end{itemize} % ends low level
% \end{frame}

@@

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
  \only<1>{\input{independent}}
  \only<2>{\input{cycle}}
  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
  holds
  \begin{enumerate}
  \item $\emptyset \in \mathscr{I}$,
  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
    \mathscr{I}$,
  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
    \setminus J$ such that $J \cup i \in \mathscr{I}$,
  \end{enumerate}
  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
  on $S$. This defines an independence relation $i \nsim_M I$ iff
  $I\cap i \in \mathscr{I}$.

@@

@@

\section{Matroids from logically dependent hypotheses}
\begin{frame}
  \frametitle{Dependent hypotheses}
  \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
    \begin{itemize}
    \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
    \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
      $\sim_H = \sim_M$?
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
\frametitle{Matroids from logically dependent hypotheses}

@@

\begin{frame}
  \frametitle{Stepwise test procedure}
  \begin{block}{Multiple testing procedure for logically related hypotheses}
    Assume that $\mathscr{H}$ with dependence relation $\sim_H$
    defines a matroid on $M$ on $S$. Then a test procedure that
    rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
    \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
    controls the FWER in the strong sense.
  \end{block}

@@

\section{Graphical approaches for logically dependent hypotheses}

@@

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
  \includegraphics[height=1\textwidth]{spheresfukuda.eps}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item 4 contrast hypotheses about 3 means
    \item Cut the hyperplanes with the sphere in $\R^3$ (makes
      visualisation easier)
    \item Cells of the sphere arrangment correspond to sign patterns
      of $\mathscr{H}$
    \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
    \end{itemize}
  \end{minipage}
\end{tabular}

@@

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outlook}
\begin{itemize}
\item MCPmod on \sout{steroids} matroids
\end{itemize}
\end{frame}

>>>>

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{ulem}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\sl i.e.},~}
\newcommand{\eg}{{\sl e.g.},~}
\newcommand{\parentd}{../}
\newcommand{\R}{\mathbb{R}}
\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage
\setbeamertemplate{blocks}[rounded][shadow=true]

\title{Matroids \& testing in families of dependent hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

@@

\section{Algebraically dependent hypotheses}

\begin{frame}
\frametitle{Some history:}
\framesubtitle{Testing hypotheses of no differences in pairwise
  comparisons of 3 treatments}

% Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):

\begin{block}{$H_1:\mu_1=\mu_2,\; H_2:\mu_2=\mu_3,\;H_3:\mu_1=\mu_3$}
  \begin{enumerate}
  \item Test $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 :
    \mu_1=\mu_2=\mu_3$ at level $\alpha$,
  \item test each $H_1,H_2,H_3$ at level $\alpha$,
  \end{enumerate}
  protects the FWER at level $\alpha$.
\end{block}
\pause
  \begin{itemize}
    \item  essentially closed test, since %; no need to test pairwise
                                %intersection hypotheses since 
        $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j,$

    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
  \end{itemize}

\begin{frame}
\frametitle{Some history}
\framesubtitle{Testing hypotheses of no difference in pairwise
    comparisons of $m>3$ treatments:}

  Modified Sequentially Rejective
  Multiple test procedures \cite{shaffer1986modified} % Example: Modified Bonferroni-Holm
      \begin{itemize}
      \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
      \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test
        remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ =
        max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are
        false.
      \end{itemize}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{tabular}{ll}

  \begin{minipage}{.5\textwidth}
    \input{full4}
  \end{minipage} 
  \begin{minipage}{.5\textwidth}
    \begin{block}{Graphical representation}
      \begin{itemize}
      \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq
        m*(m-1)/2$

      \item Undirected graph %$\mathscr{G} = (V,E)$:
        \begin{itemize}
        \item vertices correspond to means, 
          %\emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
        \item edges correspond to hypotheses, 
          % \emph{i.e.}, $E = \left\{e_1,...,e_n\} \right\}$, $e_\ell = \{\mu_i,\mu_j\}$, $i
          % \neq j$, and $e_\ell \neq e_k$ for $\ell \neq k$
        \end{itemize}
      \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
      \item $I$ induces a subgraph %$\mathscr{G}_I$ 
        on the edge set
        %$E_I=\{e_i: i \in I\}$
      \end{itemize} % ends low level
    \end{block}
  \end{minipage}

\end{tabular}
\end{frame}

% \begin{block}{Example: $m=4, n=6$ :}
%       \begin{itemize}
%       \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
%       \item Dynamic: $t^*_j \le t_j$, hence more powerful than
%         "static"
%       \end{itemize}
%   \end{block}

\begin{frame}
\frametitle{Example: Modified Bonferroni Holm \cite{shaffer1986modified}}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}

    \input{shaffer}
\end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{block}{$m=4, n=6, \only<2->{j=2}$}
      % \begin{itemize}
      % \item "Static": $t_j$ = max number of true hypotheses if $j$ are
      %   false.
      % \item "Dynamic": $t^*_j $ = max. number of true hypotheses if
      %   $H_{i_1},...,H_{i_j}$ are false.
%      \item Example:  \only<3->{$ j = 2$ :}
        \begin{description}
        \item<3->[Static:]  $H_1$ and $H_2$ are false
            $\Rightarrow$ $t_2 = 3$ (at most $H_4, H_5 $ and $H_6$ can be
            simultaneously true)
        \item<7->[Dynamic:] $H_1$ and $H_6$ are false $\Rightarrow$
          $t^*_2 = 2 < t_2 =3$ (only $H_3$, $H_4$, or $H_2$, $H_5$ can be
            simultaneously true)
        \end{description}

 %     \end{itemize} % ends low level
    \end{block}
  \end{minipage}
\end{tabular}
\end{frame}

% \begin{frame}
% \frametitle{Pairwise mean comparisons (two-sided)}
% \begin{itemize}

% \item Consider $\mu_i, i \in \{1,...,m\}$

% \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

% \item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
% \begin{itemize}
% \item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
% \item edges correspond to hypotheses, \emph{i.e.}, $E =
%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
%      j$, and $e_l \neq e_k$ for $l \neq k$
% \end{itemize}
% \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
% \item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
%  \in I\}$
% \end{itemize} % ends low level
% \end{frame}

% \begin{frame}
% \frametitle{Example 1: 4 Groups}
% \input{full4ext}
% \end{frame}

@@

\begin{tabular}{ll}
 \begin{minipage}{.44\textwidth}
   \begin{block}{\footnotesize Pairwise comparisons of 4 means}
     {\footnotesize
       \begin{displaymath}
         \bs{C}^t = \left(
           \begin{array}{rrrr}
             -1 & 1 & 0 & 0 \\
             -1 & 0 & 1 & 0 \\
             -1 & 0 & 0 & 1 \\
             0 & -1 & 1 & 0 \\
             0 & -1 & 0 & 1 \\
             0 & 0 & -1 & 1
           \end{array}
         \right)
       \end{displaymath}}
   \end{block}
 \end{minipage} &
 \begin{minipage}{.56\textwidth}
\pause
   \begin{block}{\footnotesize Modified Williams Contrasts for Dose Response \cite{bretz2010multiple}}
      {\footnotesize
       \begin{displaymath}
         \bs{C}^t=\left(
           \begin{array}{rrrr}
             1 & 0 & 0 & -1\\
             1 & 0 &-.51 & -.49 \\
             1 & -.35 & -.33 & -.32\\
             .51 & .49 & -.51 & -.48\\
             .51 & .49 & 0 &  -1\\
             .35 & .33 & .32& -1
           \end{array}\right)
       \end{displaymath}}
   \end{block}
 \end{minipage}
\end{tabular}

@@

\begin{itemize}
\item $H_i$ depends on $H_I$: $H_{I \cup i} = H_{I}$
\item induces a dependence relation $i \sim_H I$ on index sets $I \in \mathcal{P}(S)$
\item $I$ independent iff for all $i
  \in I$ $i \nsim_H I\setminus i$ (\ie $H_{I} \subsetneq H_{I\setminus i}$)
\pause
\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
 column rank
\end{itemize} % ends low level
\end{frame}

@@

\item Matroids generalize the structures of graphs and vector spaces,
 \eg cycles, bases, independent sets
\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
 structure
\item {\em Cryptomorphism} \ie many ways to define a matroid:
 \begin{itemize}
 \item From a graph perspective: Circuit axioms.
 \item From a vector space perspective: Independence axioms.
 \end{itemize}
%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
\end{itemize} % ends low level
\end{frame}

@@

\begin{tabular}{ll}
 \begin{minipage}{.5\textwidth}
   \begin{block}{Circuits of a matroid $M$ on $S$:}
       A collection $\mathscr{C}$ of subsets of $S$
       for which holds
     \begin{enumerate}
     \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
     \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$
       then $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
     \end{enumerate}
   \end{block}

 \end{minipage} &

 \begin{minipage}{.5\textwidth}
   \only<2->{\input{cycle}}
 \end{minipage} 
\end{tabular}

@@

@@

\begin{tabular}{ll}
 \begin{minipage}{.5\textwidth}
   \begin{block}{Independent sets of a matroid $M$ on $S$}
     A a collection $\mathscr{I}$ of subsets of $S$ for which
     holds
     \begin{enumerate}
     \item $\emptyset \in \mathscr{I}$,
     \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
       \mathscr{I}$,
     \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in
       I \setminus J$ such that $J \cup i \in \mathscr{I}$,
     \end{enumerate}
   \end{block}
 \end{minipage} &

 \begin{minipage}{.5\textwidth}
 \only<1>{\input{independent}}
 \only<2>{\input{cycle}}
 \invisible<1>{\alert{\footnotesize minimally dependent set!}}
 \end{minipage} 
\end{tabular}

@@

\begin{frame}
  \frametitle{}
  \begin{itemize}
  \item Circuits are minimally dependent sets (remove any element to
    get an independent set)
  \item Independent sets are sets for which no subset of elements is a
    circuit
  \item Maximally independent sets are bases (Base axioms)
  \item Cardinality of largest independent subset is the rank of a set
    (Rank axioms)
  \item Maximally dependent sets are called flats (Closure axioms)
  \end{itemize}
\end{frame}

@@

\section{Matroids from algebraically dependent hypotheses}
\begin{frame}
 \frametitle{Dependent hypotheses}
 \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
   \begin{itemize}
   \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
   \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
     $\sim_H = \sim_M$?
   \end{itemize}
 \end{block}
\end{frame}

\begin{frame}
\frametitle{Matroids from algebraically dependent hypotheses}

@@

\begin{frame}
 \frametitle{Stepwise test procedure}
 \begin{block}{Multiple testing procedure for algebraically related hypotheses}
   Assume that $\mathscr{H}$ with dependence relation $\sim_H$
   defines a matroid on $M$ on $S$. Then a test procedure that
   rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
   \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
   controls the FWER in the strong sense.
 \end{block}

@@

\section{Graphical approaches for algebraically dependent and
  contextually related hypotheses}

@@

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
 \includegraphics[height=1\textwidth]{spheresfukuda.pdf}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item 4 contrast hypotheses about 3 means
   \item Cut the hyperplanes with the sphere in $\R^3$ (makes
     visualisation easier)
   \item Cells of the sphere arrangment correspond to sign patterns
     of $\mathscr{H}$
   \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
   \end{itemize}
 \end{minipage}
\end{tabular}

@@

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
\item Efficient algorithms from the realms of combinatorics and algebraic geometry
\item {\em Scum principle} everything important happens at the top
\item Stepwise procedure for algebraically dependent and contextually
  related hypotheses (graphical approach)
\item Promising results also for one-sided hypotheses
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outlook}
\begin{itemize}
\item Inclusion of procedures for two-sided tests in \texttt{gMCP} 
\item Implementation of one-sided problems using polyhedral
  programming libraries
\item Potential applications:
  \begin{itemize}
  \item \texttt{MCPmod} on \sout{steroids} matroids
  \item Model selection
  \item Logic regression
  \end{itemize}
\item Tests for intersection hypotheses based on maximally independent
  sets 
\end{itemize}
\end{frame}
